---
title: "Hydrocron Example Lake Data Download"
author: "Mike Talbot"
date: "`r Sys.Date()`"
output: pdf_document
---

This notebook downloads and parses the lake level time series data in the Hydrocron example: https://podaac.github.io/hydrocron/examples.html

It's basically a minimal example of how the API works and how the JSON can be parsed in R. 

You can view the lakes data at the link below. The relevant information needed for the API request is simply the Lake ID (replace the `lake_id` variable in the script below). You can also download data directly through the viewer.

https://apps.usgs.gov/wisp?lat=40.579328&lon=-105.095101&zoom=10.89&basemap=%22Light%22&useDevStyles=false&activeLayers=%5B%7B%22id%22%3A%22lakes%22%7D%5D

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(httr)
library(jsonlite)
```

# Download
```{r download_example_lake_data}
# Define download directory and file name
out_dir <- "data"

# Define an output file name
out_file <- "example_data.csv"

# Define the ID of the lake you want to query
lake_id <- "7420101733"

# Define the start and end times of the time series of lake levels
start_time <- "2023-01-31T00:00:00Z"
end_time <- "2025-01-31T00:00:00Z"

# Define the example API URL
url <- str_glue("https://soto.podaac.earthdatacloud.nasa.gov/hydrocron/v1/timeseries?feature=PriorLake&feature_id={lake_id}&start_time={start_time}&end_time={end_time}&fields=lake_id,time_str,wse,area_total,quality_f,collection_shortname,crid,PLD_version,range_start_time&output=csv")

# Make the GET request to the API
response <- GET(url)

# Check the status of the response
if (status_code(response) == 200) {
  # Parse the JSON response
  json_content <- content(response, as = "text", encoding = "UTF-8")
  parsed_json <- fromJSON(json_content)
  
  # Extract the CSV data from the JSON
  csv_content <- parsed_json$results$csv
  
  # Write the extracted data to a CSV file
  write_file(csv_content, file.path(out_dir, out_file))
  cat("Data downloaded successfully.\n")
} else {
  cat("Failed to download data. Status code:", status_code(response), "\n")
}
```

\newpage
# Plot
```{r plot_example_lake_data}
# Read data
data <- read_csv(file.path(out_dir, out_file), show_col_types = F)

# View first six rows ("head") of data
print(head(data))

# Filter out "No Data" (really large or small) values
data <- data %>%
  filter(wse > -1E10 & wse < 1E10)

# Plot data
ggplot(data, aes(x = time_str, y = wse)) +
  geom_point(col = "blue") +
  geom_line(col = "blue") + 
  xlab("Date/Time") +
  ylab("Water Surface Elevation (m)") +
  ggtitle("Example Lake Data from Hydrocron") +
  theme_bw()
```